---
title: "MLST Lepto - Gene"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE, message = FALSE}
library(tidyverse)
library(knitr)
#import data
#cov <- read.csv('~/Projects/bactocap/datasets/mlst/analysis/mlst-geneonly-cov.csv', col.names = c("sample_name", "target_name", "target_len", "reads_mapped", "unmapped_frags"))
#cov <- cov %>% extract(target_name, into = c("sample_pref", "target_org"), "_")

cov <- read.csv('~/Projects/bactocap/datasets/mlst/analysis/mlst-geneonly-cov.csv', col.names = c("sample_name", "target_name", "target_len", "reads_mapped", "unmapped_frags"))
cov <- cov %>% separate(target_name, into  = c('target_gene', 'target_org', 'target_type'))
```

Crude per-target depth-of-coverage calculated by multiplying the number of reads mapped to each target by read length (75) and then dividing by the target length. Summary stats tabulated below.

**Table 1: Summary stats for per-target per-sample d.o.c**
```{r meandoc, warning = FALSE, echo =FALSE, message = FALSE}
doc <- cov %>% 
  filter(target_org == 'Lepto') %>% 
  mutate(meandoc = (reads_mapped * 75)/target_len) %>% select(sample_name, target_gene, meandoc) %>% group_by(sample_name) %>% 
  summarise(mean = mean(meandoc), median = median(meandoc), Max = max(meandoc), Min = min(meandoc))
kable(doc)
  

cov %>% 
  filter(target_org == 'Lepto') %>% 
  mutate(meandoc = (reads_mapped * 75)/target_len) %>% select(sample_name, meandoc) %>% 
  ggplot(aes(x=sample_name, y = meandoc)) +
  xlab("Mean depth-of-coverage") +
  xlab("Sample") +
  geom_boxplot() +
  geom_jitter() +
  theme_minimal()

doc <- cov %>% 
  filter(target_org == 'Lepto') %>% 
  mutate(meandoc = (reads_mapped * 75)/target_len) %>% select(sample_name, meandoc)

```


**Figure 1: Jittered boxplot for mean d.o.c per-sample per-target.** A plot of the raw data (mean per-sample, per-target d.o.c)

**Table 2: Mapping summary statistics** Per-sample number of mapped reads, mapped reads as a fraction of the total readset, and fraction of the per-sample readset mapping to 'Lepto' targets
```{r mapreads, warning = FALSE, echo =FALSE, message = FALSE}
t <- cov %>% 
  select(sample_name, reads_mapped, unmapped_frags) %>%
  group_by(sample_name) %>% 
  summarise(Mapped_Reads = sum(reads_mapped), Unmapped_Reads = sum(unmapped_frags)) %>% 
  mutate(Fraction_Of_Reads_Mapped = Mapped_Reads / (Mapped_Reads + Unmapped_Reads)) %>% 
  select(sample_name, Fraction_Of_Reads_Mapped, Mapped_Reads)

#create larger table with breakdown of target/offtarget
tib <- cov %>% mutate(.,
               target = case_when(
                 target_org == 'Lepto' ~ 'On-target_reads',
                 target_org != 'Lepto' ~ 'Off-target_reads',
                 TRUE ~ NA_character_)) %>% 
  select(sample_name, target, reads_mapped) %>% 
  group_by(sample_name, target) %>% 
  summarise(reads = sum(reads_mapped)) 

#create larger table with breakdown of target/offtarget
tab <- tib %>% pivot_wider(names_from = target, values_from = reads) %>% 
  mutate(Fraction_On_Target = `On-target_reads`/(`On-target_reads`+`Off-target_reads`)) %>% 
  select(sample_name, `Off-target_reads`, `On-target_reads`, Fraction_On_Target)

tob <- cbind(t$sample_name, t$Mapped_Reads, t$Fraction_Of_Reads_Mapped, tab$Fraction_On_Target)


colnames(tob) <- c("Sample Name", "Mapped Reads", "Fraction Of Reads Mapped", "Fraction of Readset On-Target")
kable(tob)
```
**Table 3: Breakdown of mean d.o.c per-target, per-sample, for the gene targets containing 'Lepto' in the sequence name**
```{r covbytarget, warning = FALSE, echo =FALSE, message = FALSE}
targ <- cov %>% filter(target_org == 'Lepto') %>% 
  mutate(meandoc = (reads_mapped * 75)/target_len) %>% select(sample_name, target_gene, meandoc) %>% 
  pivot_wider(names_from = sample_name, values_from = meandoc)
kable(targ)
```
